R for Clinical Study Reports and Submission

### 2021-09-17 tlf-overview.Rmd
Improve flow and fix typos

In clinical trial development, it is a critical step to 
submit information on medical products to regulatory agencies. 
[Electronic Common Technical Document (eCTD)](https://en.wikipedia.org/wiki/Electronic_common_technical_document) 
has become a worldwide regulatory submission standard format. 
For example, [the United States Food and Drug Administration (US FDA) required new drug applications, biologics license applications must be submitted using eCTD format](https://www.fda.gov/drugs/electronic-regulatory-submission-and-review/electronic-common-technical-document-ectd). Clinical Data Interchange Standards Consortium (CDISC) provided a 
[pilot project following ICH E3 guidance](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse).

Within eCTD, clinical study reports (CSR) are located at module 5. 
[ICH E3 guidance](https://www.ich.org/page/efficacy-guidelines) provides 
compilation of structure and content of clinical study reports.

A typical CSR contain full details of an individual clinical study 
including clinical and statistical description, presentations, and analyses. 
A large number of tables and figures are incorporated into the main text and appendices of a CSR. 
In CDISC pilot project, an 
[example CSR](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/53-clin-stud-rep/535-rep-effic-safety-stud/5351-stud-rep-contr/cdiscpilot01/cdiscpilot01.pdf) 
is also provided. If you are interested in more examples of clinical study reports, 
the European Medicines Agency (EMA) publishes clinical data including 
clinical study reports submitted by pharmaceutical companies 
on [EMA clinical data website](https://clinicaldata.ema.europa.eu/web/cdp/home).

Building CSR is teamwork with clinicians, medical writers, statisticians, and statistical programmers. 
Here, we focus on the work and deliverables completed by statisticians and statistical programmers. 
In an organization, a group of statisticians and statistical programmers commonly work together to 
define, develop, validate and deliver tables, listings, and figures (TLFs) required for a CSR.
Microsoft Word is widely used to prepare CSR in the pharmaceutical industry. 
Therefore, `rtf`, `doc`, `docx` are commonly used formats to deliver TLFs.

In this chapter, our focus is to illustrate how to create tables, listings, and figures (TLFs) in the RTF format 
that is commonly used in a CSR. The examples are in compliance of 
[FDA's Portable Document Format (PDF) Specifications](https://www.fda.gov/media/76797/download)

@@

These TLFs are typically used to summarize efficacy and safety 
statistical analysis results of a drug (or treatment).

@@

- Section 10: Study patients 
- Section 11: Efficacy evaluation
- Section 12: Safety evaluation 
- Section 14: ables, Figures, and Graphs referred to but not included in the text
- Section 16: Appendices

@@

We provided a brief introduction of `r2rtf` and show how to transfer 
data frames into Table, Listing, and Figure (TLFs).

>>>>

In clinical trials, a critical step is
submission to regulatory agencies. 
[Electronic Common Technical Document (eCTD)](https://en.wikipedia.org/wiki/Electronic_common_technical_document) 
has become a worldwide regulatory submission standard format. 
For example, [the United States Food and Drug Administration (US FDA) required new drug applications, biologics license applications must be submitted using eCTD format](https://www.fda.gov/drugs/electronic-regulatory-submission-and-review/electronic-common-technical-document-ectd). Clinical Data Interchange Standards Consortium (CDISC) provided a 
[pilot project following ICH E3 guidance](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse).

Within eCTD, clinical study reports (CSRs) are located at module 5. 
[ICH E3 guidance](https://www.ich.org/page/efficacy-guidelines) provides 
compilation of structure and content of clinical study reports.

A typical CSR contains full details on the methods and results of a individual clinical study.
In support of the statistical anlaysis, a large number of tables, listings and figures are
incorporated into the main text and appendices.
In CDISC pilot project, an 
[example CSR](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/53-clin-stud-rep/535-rep-effic-safety-stud/5351-stud-rep-contr/cdiscpilot01/cdiscpilot01.pdf) 
is also provided. If you are interested in more examples of clinical study reports, 
the European Medicines Agency (EMA) publishes clinical data including 
clinical study reports submitted by pharmaceutical companies 
on [EMA clinical data website](https://clinicaldata.ema.europa.eu/web/cdp/home).

Building CSRs is a teamwork between clinicians, medical writers, statisticians, statistical programmers
and other relevant specialists such as experts on biomarkers. 
Here, we focus on the work and deliverables completed by statisticians and statistical programmers. 
In an organization, they commonly work together to 
define, develop, validate and deliver tables, listings, and figures (TLFs) required for a CSR to
summarize the safety and/or safety of the pharmaceutical product.
Microsoft Word is widely used to prepare CSR in the pharmaceutical industry. 
Therefore, `rtf`, `doc`, `docx` are commonly used formats in their deliverables.

In this chapter, our focus is to illustrate how to create tables, listings, and figures (TLFs) in RTF format
that is commonly used in a CSR. The examples are in compliance of 
[FDA's Portable Document Format (PDF) Specifications](https://www.fda.gov/media/76797/download)

@@

@@

- Section 10: Study patients 
- Section 11: Efficacy evaluation
- Section 12: Safety evaluation 
- Section 14: Tables, Listings and Figures referrals but not included in the text
- Section 16: Appendices

@@

We provided a brief introduction of `r2rtf` and show how to transfer 
data frames into Table, Listing, and Figures (TLFs).
